Regression

### 2015-09-04 Regression.rmd
further work on regression.rmd

final.output<- function(df=all7d, samp = 800) {
        #create sample df
        sample = df[samp,]
        sample = sample[,!is.na(sample)]
        vars= names(sample)
        vars = vars[! vars %in% c('mlsto','Year','inv.mlsto','timespan','client.totinv','code.director','Num.disc',
                                  'Biz.type','Post.Code','code.ProjEng','return.pdol','cost.mlsto','hrs.mlsto',
                                  'balance.mlsto','client.totinv.log','pc.majpos')]
        variables = vars

        print(
                list("sample case", t(sample %>% select(-Year, -timespan.cbrt, -inv.mlsto.log, -client.totinv.log, -pc.majpos.log)))
        )

        #make reduced dataset with complete cases for all variables we want to use
        reduced = df[complete.cases(df[,variables]),]

@@

        final.RMSE= RMSE(ans$answer, ans$final.lm.fit) %>% round(4)

@@

        print(
                paste("Invoiced= $", sample$inv.mlsto, ",   predicted cost= $", predict.cost %>% round(1), sep="")
        )

        print(
                paste("Actual cost= $", sample$cost.mlsto, sep="")
        )

@@

final.output(df=all7d, samp = 500)

@@

>>>>

final.output<- function(df=all7d, samp = 600) {
        #create sample df
        sample = df[samp,]
        sample = sample[,!is.na(sample)]
        vars= names(sample)
        vars = vars[! vars %in% c('mlsto','Year','inv.mlsto','timespan','client.totinv','code.director','Num.disc',
                                  'Biz.type','Post.Code','code.ProjEng','return.pdol','cost.mlsto','hrs.mlsto',
                                  'balance.mlsto','client.totinv.log','pc.majpos')]
        variables = vars

        print(
                list("sample case", t(sample %>% select(-Year, -timespan.cbrt, -inv.mlsto.log, -client.totinv.log)))
        )

        #make reduced dataset with complete cases for all variables we want to use
        reduced = df[-samp,]
        reduced = reduced[complete.cases(reduced[,variables]),]

@@

        #random RMSE
        r= mean(reduced$return.pdol)
        random.RMSE= RMSE(reduced$return.pdol, r, print=T) %>% round(3)

        print(
        paste("Random Guess RMSE=", random.RMSE , sep= " ")
        )

        #final RMSE for all rows in reduced in a regression
        final.RMSE= RMSE(ans$answer, ans$final.lm.fit) %>% round(3)

@@

        print(
                paste("Invoiced= $", sample$inv.mlsto %>% round(2), ",   predicted cost= $", predict.cost %>% round(1), sep="")
        )

        print(
                paste("Actual cost= $", sample$cost.mlsto %>% round(2), sep="")
        )

@@

final.output(df=all7d, samp = 800)

@@

Check what RMSE would be for random guess

```{r, echo=F}

m<- mean(all7d$return.pdol)

RMSE<- function(answer, guess, print= TRUE){
        error= answer-guess
        root.error= sqrt(mean(error^2))
        if(print == TRUE){
        print(root.error)
        }
}

RMSE(all7d$return.pdol, m, print=T)

RMSE from random guess is 53 cents.

# delete any code.client or contact. client which has less than 3 entries!
# need to revisit how i got my inv.mlsto, balance.mlsto, cost.mlsto for job 2013.301.2

### 2015-09-04 rand_forest.rmd
further work on regression.rmd

# 
# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
# opts_knit$set(root.dir= "C:/Users/n9232371/Documents/Consultbusiness/data")
all7d<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7d.csv')[,-1]
all7d<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all7d.csv')[,-1]
all7d$Post.Code<- as.factor(all7d$Post.Code)

@@

train_ind <- sample(seq_len(nrow(all7d)), size = nrow(all7d)*2/3)

train <- all7d[train_ind, ]
test <- all7d[-train_ind, ]

#create tree with training data only
tree.all7d<- rpart(return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type + code.ProjEng, data=train, control = rpart.control(cp= 0.001))
# summary(tree.all7d)

#have a look at cp values, automatic cross validation in rpart
printcp(tree.all7d)
plotcp(tree.all7d)
#min cp = 0.013 will give simplest tree within 1 std of x-val rel error
tree.all7d<- rpart(return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type + code.ProjEng, data=train,
                  control = rpart.control(cp= 0.013))

@@

prp(tree.all7d, extra = 1, split.fun= split.fun, tweak=1, faclen =5, branch.type= 5)
#make prediction with current tree
tree.pred<-predict(tree.all7d, test)
#calculate RMSE for test data
RMSE(test$return.pdol, tree.pred)
# 49.8 cents, already close to regression..

@@

heat.tree(tree.all7d, extra = 1, split.fun= split.fun, tweak=1, faclen =6, branch.type= 5)

>>>>

# 
# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
# opts_knit$set(root.dir= "C:/Users/n9232371/Documents/Consultbusiness/data")
all7d.tree<- read.csv('C:/Users/n9232371/Documents/Consultbusiness/data/all7d.csv')[,-1]
all7d.tree<- read.csv('~/OneDrive/shared files/Bligh Tanner/masters/data/all7d.csv')[,-1]
all7d.tree$Post.Code<- as.factor(all7d$Post.Code)

@@

train_ind <- sample(seq_len(nrow(all7d.tree)), size = nrow(all7d.tree)*2/3)

train <- all7d.tree[train_ind, ]
test <- all7d.tree[-train_ind, ]

#create tree with training data only
tree.all7d.tree<- rpart(return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type + code.ProjEng, data=train, control = rpart.control(cp= 0.001))
# summary(tree.all7d.tree)

#have a look at cp values, automatic cross validation in rpart
printcp(tree.all7d.tree)
plotcp(tree.all7d.tree)
#min cp = 0.013 will give simplest tree within 1 std of x-val rel error
tree.all7d.tree<- rpart(return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type + code.ProjEng, data=train,
                  control = rpart.control(cp= 0.013))

@@

prp(tree.all7d.tree, extra = 1, split.fun= split.fun, tweak=1, faclen =5, branch.type= 5)
#make prediction with current tree
tree.pred<-predict(tree.all7d.tree, test)
#calculate RMSE for test data
RMSE(test$return.pdol, tree.pred)
# 49.8 cents, already close to regression..

@@

heat.tree(tree.all7d.tree, extra = 1, split.fun= split.fun, tweak=1, faclen =6, branch.type= 5)
